The error in Bayesian phylogenetic reconstruction when speciation co-occurs

### 2018-12-03 article.tex
Put hypotheses in figures

% Hypothesis: general
\item We hypothesize that the error made today, using BD tree priors,
increases with an increased number or stronger effect of multiple birth events.
This is straightforward: without multiple birth events or such event having
no effect, the MBD model falls back to a BD model. We expect larger
errors when we deviate more from the BD model's assumptions.

\item Additionally, we hypothesize MBD having a stronger effect if the normal
speciation process is less pronounced. The more speciations are caused by the
BD process, there are relatively less multiple-birth events.

\item We expect the effect of extinction rates to be neutral, as extinctions will
hit lineages created by both speciation processes equally.

\item To put into an explicit equation, 
we expect the error made be correlated to the number 
of species created by the multiple-birth process over the total number 
of species created:

@@

\item Due to the proportionality of the term within $f$, 
we expect the timing of a multiple births event (be it close to the
crown age or close to the present) to have no effect. Compared to a late
multiple birth event, an early multiple birth event may have a 
longer-lasting effect (as the next speciation event will be later), but it
will create less new species, as there are still fewer taxa.

@@

\item We start simulating $N_{S} = 1000$ MBD trees. 
From each MBD tree, a DNA sequence alignment is simulated. 
For each sequence alignment we then perform a Bayesian analysis 
to recover a posterior distribution of trees, 
each composed of $N_{P}$ phylogenies. 
Such analysis is performed using 
the 'pirouette' package (\cite{pirouette}) to call the BEAST2 tool 
suite from R. 
We let the Bayesian analysis assume a BD prior in both cases, 
to investigate the extent of the error we make under this assumption.

@@

\item We then exploit such parameters to generate a BD tree 
using the function "\texttt{tess.sim.taxa.age}" 
from the package "\texttt{TESS}" (\cite{Hoehna2013}). 
We simulate the tree in such a way the new tree 
has the same number of tips and the same crown age as the MBD tree. 
We furthermore require that the BD tree conserve the topology of the MBD tree.

>>>>

% Hypothesis: general
\item We hypothesize that the error made today, using BD tree priors,
increases with an increased number or stronger effect of multiple birth events.
This is straightforward: without multiple birth events or such event having
no effect, the MBD model falls back to a BD model. We expect larger
errors when we deviate more from the BD model's assumptions. 
Additionally, we hypothesize MBD having a stronger effect if the normal
speciation process is less pronounced. The more speciations are caused by the
BD process, there are relatively less multiple-birth events.
To put this hypothesis, H1, into an explicit equation, 
we expect the error made be correlated to the number 
of species created by the multiple-birth process over the total number 
of species created:

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_h_1.png}
  \caption{
    Hypothesis 1: we expect the error made be correlated to the number 
    of species created by the multiple-birth process over the total number 
    of species created
  }
  \label{fig_h_1}
\end{figure}

\item We have the hypothesis, H2, 
that the effect of extinction rates is neutral, 
as extinctions will hit lineages created by both speciation processes equally.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_h_2.png}
  \caption{
    Hypothesis 2: the effect of extinction rates is neutral
  }
  \label{fig_h_2}
\end{figure}

\item Due to the proportionality of the term within $f$, 
we have the hypothesis, H3, 
that the timing of a multiple birth event (be it close to the
crown age or close to the present) to have no effect. Compared to a late
multiple birth event, an early multiple birth event may have a 
longer-lasting effect (as the next speciation event will be later), but it
will create less new species, as there are still fewer taxa.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_h_3.png}
  \caption{
    Hypothesis 3: the timing of a multiple birth event has no effect.
  }
  \label{fig_h_3}
\end{figure}

@@

\item We start simulating $N_{S} = 1000$ MBD trees,
with either 50, 100 and 200 taxa.
We have the hypothesis H5, that the number of taxa does not
have an effect on the error being made, as there is no
diversity dependency in any of the processes.
We have the hypothesis H6, that for a higher number of taxa 
the variance in the error decreases, as more information is present
in the simulated phylogenies.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_h_5.png}
  \caption{
    Hypothesis 5: the number of taxa does not
    have an effect on the error being made
  }
  \label{fig_h_5}
\end{figure}

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_h_6.png}
  \caption{
    Hypothesis 6: for a higher number of taxa 
    the variance in the error decreases
  }
  \label{fig_h_6}
\end{figure}

\item From each MBD tree, a DNA sequence alignment is simulated. 
For each sequence alignment we then perform a Bayesian analysis 
to recover a posterior distribution of trees, 
each composed of $N_{P}$ phylogenies. 
Such analysis is performed using 
the 'pirouette' package (\cite{pirouette}) to call the BEAST2 tool 
suite from R. 
We let the Bayesian analysis assume a BD prior in both cases, 
to investigate the extent of the error we make under this assumption.

@@

\item We then exploit such parameters to generate a BD tree 
using the function "\texttt{tess.sim.taxa.age}" 
from the package "\texttt{TESS}" (\cite{Hoehna2013}). 
We simulate the tree in such a way the new tree 
has the same number of tips and the same crown age as the MBD tree. 
We furthermore require that the BD tree conserve the topology of the MBD tree.
We have hypothesis H4 that, compared to the MBD trees, 
the error will be less in the BD twin tree.
The difference between the errors made in MBD and twin BD trees indicates
the impact the MBD process has on the error we make in inference using a
contemporary BD prior.

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{fig_h_4.png}
  \caption{
    Hypothesis 4: compared to the MBD trees, 
    the error will be less in the BD twin tree
  }
  \label{fig_h_4}
\end{figure}
